Trends of SDR in Thessaloniki

### 2023-11-05 MDPI_submition.Rmd
Commit 2023-11-05 20:14

# front matter
simplesummary: |
  A Simple summary goes here.

@@

# back matter
keywords: |
  GHI; SDR; Solar Brigthening/Dimming.
## list three to ten pertinent keywords specific to the article, yet reasonably common within the subject discipline

@@

We have demonstrated that in the period
`r paste(year(min(ALL_1_daily_DESEAS$Date)), " - ", year(max(ALL_1_daily_DESEAS$Date)))`, there is positive trend in SDR of for all-sky conditions ($`r alltrend`\,\%/y$).
A previous study [@Bais2013] for the period 1993 - 2011 the trend was found to be $0.33\,\%/y$.
The increase of this trend, indicates that the phenomenon and probably the causes, are still occurring.
Accordingly, we found a similar trend for clear-sky conditions ($`r cleartrend`\,\%/y$) that further supports that the brightening can be independent of the cloud occurrences, although, cloud formation can also attenuate SDR.
We have calculated the effect of TSI variability to be a negligible factor in SDR variability.
Unfortunately, there is no available data for this period for the aerosols, in order to directly estimate their effect on SDR, which is the most probable main cause.
The attenuation of SDR by aerosols over Europe have been proposed as major factor by @Wild2021.
The dimming effect on SDR under cloudy-sky conditions ($`r cloudtrend`\,\%/y$),suggests that cloud optical thickness is decreasing during this period.
Because we have no adequate data to investigate the long term changes of cloud thickness in the region, we can not verify if the negative SDR trend we observe under under cloudy-skies can be attributed solely to changes in clouds.

>>>>

# front matter
# simplesummary: |
#   A Simple summary goes here.

@@

# back matter
keywords: |
  GHI; SDR; Solar Brigthening/Dimming, aerosols, clouds.
## list three to ten pertinent keywords specific to the article, yet reasonably common within the subject discipline

@@

We have demonstrated that in the period
`r paste(year(min(ALL_1_daily_DESEAS$Date)), " - ", year(max(ALL_1_daily_DESEAS$Date)))`, there is a positive trend in SDR of ($`r alltrend`\,\%/y$) (brightening) (positive trend) in Thessaloniki, Greece, under all-sky conditions.
A previous study [@Bais2013] for the period 1993 - 2011 found also a positive trend of $0.33\,\%/y$.
The increase of this trend indicates that the brightening of SDR continues and is probably caused by continuing decreases in aerosol optical depth and the optical thickness of clouds over the area.
Moreover, we found a similar trend under clear-sky conditions ($`r cleartrend`\,\%/y$) that further supports the assumption that the brightening is caused mainly by decreasing aerosols.
Unfortunately, for the entire period there is no available data for the aerosols, in order to quantify their effect on SDR. 
However, @Siomos2020 have shown that aerosol optical depth over Thessaloniki is decreasing constantly at least up to 2018.
The attenuation of SDR by aerosols over Europe have been proposed as major factor by @Wild2021.
The dimming effect on SDR under cloudy-sky conditions ($`r cloudtrend`\,\%/y$), suggests that cloud optical thickness is decreasing during this period.
Because we have no adequate data to investigate the long term changes of cloud thickness in the region, we cannot verify if the negative SDR trend we observe under under cloudy-skies can be attributed solely to changes in clouds.

### 2023-11-05 MDPI_submition.tex
Commit 2023-11-05 20:14

% Simple summary
\simplesumm{A Simple summary goes here.}

@@

% Keywords
\keyword{GHI; SDR; Solar Brigthening/Dimming.}

@@

We have demonstrated that in the period 1993 - 2023, there is positive
trend in SDR of for all-sky conditions (\(0.38\,\%/y\)). A previous
study \citep{Bais2013} for the period 1993 - 2011 the trend was found to
be \(0.33\,\%/y\). The increase of this trend, indicates that the
phenomenon and probably the causes, are still occurring. Accordingly, we
found a similar trend for clear-sky conditions (\(0.35\,\%/y\)) that
further supports that the brightening can be independent of the cloud
occurrences, although, cloud formation can also attenuate SDR. We have
calculated the effect of TSI variability to be a negligible factor in
SDR variability. Unfortunately, there is no available data for this
period for the aerosols, in order to directly estimate their effect on
SDR, which is the most probable main cause. The attenuation of SDR by
aerosols over Europe have been proposed as major factor by
\citet{Wild2021}. The dimming effect on SDR under cloudy-sky conditions
(\(-0.28\,\%/y\)),suggests that cloud optical thickness is decreasing
during this period. Because we have no adequate data to investigate the
long term changes of cloud thickness in the region, we can not verify if
the negative SDR trend we observe under under cloudy-skies can be
attributed solely to changes in clouds.

>>>>

% Simple summary

@@

% Keywords
\keyword{GHI; SDR; Solar Brigthening/Dimming, aerosols, clouds.}

@@

We have demonstrated that in the period 1993 - 2023, there is a positive
trend in SDR of (\(0.38\,\%/y\)) (brightening) (positive trend) in
Thessaloniki, Greece, under all-sky conditions. A previous study
\citep{Bais2013} for the period 1993 - 2011 found also a positive trend
of \(0.33\,\%/y\). The increase of this trend indicates that the
brightening of SDR continues and is probably caused by continuing
decreases in aerosol optical depth and the optical thickness of clouds
over the area. Moreover, we found a similar trend under clear-sky
conditions (\(0.35\,\%/y\)) that further supports the assumption that
the brightening is caused mainly by decreasing aerosols. Unfortunately,
for the entire period there is no available data for the aerosols, in
order to quantify their effect on SDR. However, \citet{Siomos2020} have
shown that aerosol optical depth over Thessaloniki is decreasing
constantly at least up to 2018. The attenuation of SDR by aerosols over
Europe have been proposed as major factor by \citet{Wild2021}. The
dimming effect on SDR under cloudy-sky conditions (\(-0.28\,\%/y\)),
suggests that cloud optical thickness is decreasing during this period.
Because we have no adequate data to investigate the long term changes of
cloud thickness in the region, we cannot verify if the negative SDR
trend we observe under under cloudy-skies can be attributed solely to
changes in clouds.
